---
title: "002 prolonged symptoms"
author: "Katie Gostic"
date: "11/1/2017"
output: html_document
---

```{r setup, include=FALSE}
setwd('~/Dropbox/R/2017_INSIGHT/')
rm(list = ls())
source('Import_FLU002.R')
cols = colors()[c(506, 499, 475)]
set.seed(91)
```
Notes on data formatting from Deborah Wentworth:

* symdur is symptom duration (days) prior to enrollment
* date info is for enrollment date
* fluvac6 and fluvac12 are flu vaccination in last 6 or last 12 months
* anyvac combines values from fluvac6 and fluvac12, whichever was collected for a given patient is reported.
* anyav is use of antivirals at any time
* anydx is any diagnosis in the medical history section
* flutype is subtype: 1=H1N1, 2=H3N2, 3=flu B, 4=PCR negative, 5=Inf A, subtype unknown, and 6=coinfection with multiple strains
* resolved is symptom resolution within 14 days
* country code is the 3 letter code for country of enrollment - think the abbreviations will be obvious


Notes on duration of symptoms:
A review by Carrat et al. (2007) of challenge studies in healthy human volunteers suggests that influenza symptoms typically resolve fully within 8-9 days of inoculation. This led us to classify patients with the following characteristics as having long-lasting symptoms:

* Symptoms lasting 10 or more days prior to study enrollment (this implies that symptoms on day 10 were still serious enough to send patients to the doctor, and most likely lasted at least a few days longer.)
* Symptoms not resolved on day 14 of enrollment (implies that symptoms lasted longer than 14 days, as most patients had symptoms for at least one day prior to enrollment.)

```{r}
# use the above definition to define a column for prolonged symptoms
dat.002$prolonged.symp = as.numeric(dat.002$symdur >= 10 | dat.002$resolved == "0")
```


#### Visualize the relationship between age and prob of H1N1 infection
```{r warning = FALSE, echo = FALSE}
# Select only cases with confirmed H1N1, and whose duration of symptoms is known
valid = dat.002[dat.002$H1N1 == 1 & !is.na(dat.002$prolonged.symp) & !is.na(dat.002$H1N1), ]
tab = table(valid$age, valid$prolonged.symp)
props = tab[,2]/rowSums(tab)
CIs = sapply(X = 1:nrow(tab), FUN = function(ii){
pp = prop.test(x = tab[ii, 2], n = sum(tab[ii,]))
pp$conf.int[1:2]
} )
plot(as.numeric(rownames(tab)), props, ylim = c(0, 1), xlim = c(10, 100), xlab = 'age', ylab = 'prop infected', col = 'lightblue', pch = 16, main = 'H1N1: Observed prop. w. prolonged symp. by age')
points(as.numeric(rownames(tab)), props, col = 'dodgerblue')
segments(x0 = as.numeric(rownames(tab)), y0 = CIs[1,], y1 = CIs[2,], col = 'lightblue')
```



#### Visualize the relationship between age and prob of H3N2 infection
```{r warning = FALSE, echo = FALSE}
valid = dat.002[dat.002$H3N2 == 1 & !is.na(dat.002$prolonged.symp) & !is.na(dat.002$H3N2), ]
tab = table(valid$age, valid$prolonged.symp)
props = tab[,2]/rowSums(tab)
CIs = sapply(X = 1:nrow(tab), FUN = function(ii){
pp = prop.test(x = tab[ii, 2], n = sum(tab[ii,]))
pp$conf.int[1:2]
} )
plot(as.numeric(rownames(tab)), props, ylim = c(0, 1), xlim = c(10, 100), xlab = 'age', ylab = 'prop infected', col = 'pink3', pch = 16, main = 'H3N2: Observed prop. prop. w. prolonged symp. by age')
points(as.numeric(rownames(tab)), props, col = 'firebrick1')
segments(x0 = as.numeric(rownames(tab)), y0 = CIs[1,], y1 = CIs[2,], col = 'pink3')
```

#### Visualize the relationship between country and prolonged symptoms

H1N1

```{r warning = FALSE, echo = FALSE}
valid = dat.002[dat.002$H1N1 == 1 & !is.na(dat.002$prolonged.symp) & !is.na(dat.002$H1N1), ]
tab = table(valid$country, valid$prolonged.symp)
props = tab[,2]/rowSums(tab)
CIs = sapply(X = 1:nrow(tab), FUN = function(ii){
pp = prop.test(x = tab[ii, 2], n = sum(tab[ii,]))
pp$conf.int[1:2]
} )
ord = order(props)
plot(1:nrow(tab), props[ord], ylim = c(0, 1), xlab = '', ylab = 'prop w. prolonged symp.', col = 'lightblue', pch = 16, main = 'H1N1: Observed prop. w. prolonged symp. by country', xaxt = 'n')
axis(1, at = 1:nrow(tab), labels = rownames(tab)[ord], las = 3)
points(1:nrow(tab), props[ord], col = 'dodgerblue')
segments(x0 = 1:nrow(tab), y0 = CIs[1,ord], y1 = CIs[2,ord], col = 'lightblue')
```

H3N2

```{r warning = FALSE, echo = FALSE}
valid = dat.002[dat.002$H1N1 == 1 & !is.na(dat.002$prolonged.symp) & !is.na(dat.002$H1N1), ]
tab = table(valid$country, valid$prolonged.symp)
props = tab[,2]/rowSums(tab)
CIs = sapply(X = 1:nrow(tab), FUN = function(ii){
pp = prop.test(x = tab[ii, 2], n = sum(tab[ii,]))
pp$conf.int[1:2]
} )
ord = order(props)
plot(1:nrow(tab), props[ord], ylim = c(0, 1), xlab = '', ylab = 'prop w. prolonged symp.', col = 'pink3', pch = 16, main = 'H3N2: Observed prop. w. prolonged symp. by country', xaxt = 'n')
axis(1, at = 1:nrow(tab), labels = rownames(tab)[ord], las = 3)
points(1:nrow(tab), props[ord], col = 'firebrick1')
segments(x0 = 1:nrow(tab), y0 = CIs[1,ord], y1 = CIs[2,ord], col = 'pink3')
```


#### Visualize the relationship between season and prob infection

H1N1

```{r warning = FALSE, echo = FALSE}
valid = dat.002[dat.002$H1N1 == 1 & !is.na(dat.002$prolonged.symp) & !is.na(dat.002$H1N1), ]
tab = table(valid$country, valid$prolonged.symp)
props = tab[,2]/rowSums(tab)
CIs = sapply(X = 1:nrow(tab), FUN = function(ii){
pp = prop.test(x = tab[ii, 2], n = sum(tab[ii,]))
pp$conf.int[1:2]
} )
#ord = order(props)
ord = 1:nrow(tab)
plot(1:nrow(tab), props[ord], ylim = c(0, 1), xlab = '', ylab = 'prop w. prolonged symp.', col = 'lightblue', pch = 16, main = 'H1N1: Observed prop. w. prolonged symp. by season', xaxt = 'n')
axis(1, at = 1:nrow(tab), labels = rownames(tab)[ord], las = 3)
points(1:nrow(tab), props[ord], col = 'dodgerblue')
segments(x0 = 1:nrow(tab), y0 = CIs[1,ord], y1 = CIs[2,ord], col = 'lightblue')
```

H3N2

```{r warning = FALSE, echo = FALSE}
valid = dat.002[dat.002$H3N2 == 1 & !is.na(dat.002$prolonged.symp) & !is.na(dat.002$H3N2), ]
tab = table(valid$season, valid$prolonged.symp)
props = tab[,2]/rowSums(tab)
CIs = sapply(X = 1:nrow(tab), FUN = function(ii){
pp = prop.test(x = tab[ii, 2], n = sum(tab[ii,]))
pp$conf.int[1:2]
} )
#ord = order(props)
ord = 1:nrow(tab)
plot(1:nrow(tab), props[ord], ylim = c(0, 1), xlab = '', ylab = 'prop w. prolonged symp.', col = 'pink3', pch = 16, main = 'H3N2: Observed prop. w. prolonged symp. by season', xaxt = 'n')
axis(1, at = 1:nrow(tab), labels = rownames(tab)[ord], las = 3)
points(1:nrow(tab), props[ord], col = 'firebrick1')
segments(x0 = 1:nrow(tab), y0 = CIs[1,ord], y1 = CIs[2,ord], col = 'pink3')
```



### H1N1 10-fold CV
We know the relationship between age and probability of infection should be non-linear. Use cross-validation to verify, and to choose the number of degrees of freedom to include in a spline
```{r fig.height=7, cache=TRUE, echo = FALSE}
# 
# ## Cross-validation template
# set.seed(23)
# library(splines)
# dat.in = dat.002[dat.002$H1N1 == 1 & !is.na(dat.002$prolonged.symp) & !is.na(dat.002$H1N1), ]
# ##  Perform 10-fold cross-validation
# # 1. Partition the data into 10 folds
# fold.bounds = matrix(  c(0:9*floor(nrow(dat.in)/10)+1, 1:10*floor(nrow(dat.in)/10)+1), nrow = 10, ncol = 2); fold.bounds[10, 2] = nrow(dat.in) # Set indices at which folds start and end
# fold.inds = sample(1:nrow(dat.in), replace = FALSE) # Randomly order the observations in the data
# 
# max.df = 12
# test.error.rate = matrix(NA, 10, max.df, dimnames = list(paste('fold', 1:10, sep = '.'), paste('df', 1:max.df, sep = '')))
# par(mfrow = c(2,2))
# 
# 
# for(ii in 1:10){
#   ## Fit the model to 9 of 10 folds
#   test.dat = dat.in[fold.inds[fold.bounds[ii,1]:fold.bounds[ii,2]], ]
#   training.dat = dat.in[fold.inds[-(fold.bounds[ii,1]:fold.bounds[ii,2])], ]
#   
#   ## Train the model
#   fits = vector('list', max.df); names(fits) = c('linear', paste('ns', 2:max.df, 'df', sep = ''))
#   fits[[1]] = glm(prolonged.symp ~ age, family = binomial, data = training.dat) # Linear fit
#   for(dd in 2:max.df){
#     fits[[dd]] = glm(prolonged.symp ~ ns(age, df = dd), family = binomial, data = training.dat) # fits a nat spline with dd df
#   }
#   
#   
# 
#   
#   ## Test the model on the test data
#   ## Simulate outcomes using predicted probs
#   ## Calculate test error rate
#   for(jj in 1:length(fits)){
#   # Predict probabilities using each fit
#   pred.probs = predict(fits[[jj]], newdata = list(age = test.dat$age), type = 'response')
#   if(ii == 8){ # For one CV itteration
#   plot.probs = predict(fits[[jj]], newdata = list(age = min(test.dat$age):max(test.dat$age)), type = 'response', se.fit = TRUE)
#   ## Visualize results
#   plot(jitter(test.dat$age, factor = 2.5), test.dat$prolonged.symp, pch = '|', xlim = c(0, 100), xlab = 'age', ylab = 'prob prolonged symp.', main = paste('df =', jj), col = 'gray')
#   lines(min(test.dat$age):max(test.dat$age), plot.probs$fit, col = 'dodgerblue')
#   }
#   ## Continue cross validation
#   sim.outcomes = sapply(X = pred.probs, FUN = function(pp) sample(c(0,1), size = 1, prob = c(1-pp, pp)))
#   test.error.rate[ii, jj] = sum(abs(sim.outcomes - test.dat$prolonged.symp))/nrow(test.dat)
# }
# }
```

Plot the overall test error rate vs. df
```{r echo = FALSE}
# par(mfrow = c(1,1))
# plot(1:max.df, colMeans(test.error.rate), ylab = 'test error rate', xlab = 'df', type = 'b', main = 'H1N1 10-fold CV to choose df')
```

Here, there is a weak preference for 4 degrees of freedom, but there's not much variation overall and the pattern is not super clear. I'm going with 4 df because a visual comparison of the fits wtih the observed patterns suggest that 4 df provides a good representation of observed patterns without a crazy number of wiggles in the spline.


### H3N2 10-fold CV
We know the relationship between age and probability of infection should be non-linear. Use cross-validation to verify, and to choose the number of degrees of freedom to include in a spline
```{r fig.height=7, cache=TRUE, echo = FALSE}
# library(splines)
# ## Cross-validation template
# dat.in = dat.002[dat.002$H3N2 == 1 & !is.na(dat.002$prolonged.symp) & !is.na(dat.002$H3N2), ]
# ##  Perform 10-fold cross-validation
# # 1. Partition the data into 10 folds
# fold.bounds = matrix(  c(0:9*floor(nrow(dat.in)/10)+1, 1:10*floor(nrow(dat.in)/10)+1), nrow = 10, ncol = 2); fold.bounds[10, 2] = nrow(dat.in) # Set indices at which folds start and end
# fold.inds = sample(1:nrow(dat.in), replace = FALSE) # Randomly order the observations in the data
# 
# max.df = 12
# test.error.rate = matrix(NA, 10, max.df, dimnames = list(paste('fold', 1:10, sep = '.'), paste('df', 1:max.df, sep = '')))
# par(mfrow = c(2,2))
# 
# 
# for(ii in 1:10){
#   ## Fit the model to 9 of 10 folds
#   test.dat = dat.in[fold.inds[fold.bounds[ii,1]:fold.bounds[ii,2]], ]
#   training.dat = dat.in[fold.inds[-(fold.bounds[ii,1]:fold.bounds[ii,2])], ]
#   
#   ## Train the model
#   fits = vector('list', max.df); names(fits) = c('linear', paste('ns', 2:max.df, 'df', sep = ''))
#   fits[[1]] = glm(prolonged.symp ~ age, family = binomial, data = training.dat) # Linear fit
#   for(dd in 2:max.df){
#     fits[[dd]] = glm(prolonged.symp ~ ns(age, df = dd), family = binomial, data = training.dat) # fits a nat spline with dd df
#   }
#   
#   
# 
#   
#   ## Test the model on the test data
#   ## Simulate outcomes using predicted probs
#   ## Calculate test error rate
#   for(jj in 1:length(fits)){
#   # Predict probabilities using each fit
#   pred.probs = predict(fits[[jj]], newdata = list(age = test.dat$age), type = 'response')
#   if(ii == 8){ # For one CV itteration
#   plot.probs = predict(fits[[jj]], newdata = list(age = min(test.dat$age):max(test.dat$age)), type = 'response', se.fit = TRUE)
#   ## Visualize results
#   plot(jitter(test.dat$age, factor = 2.5), test.dat$prolonged.symp, pch = '|', xlim = c(0, 100), xlab = 'age', ylab = 'prob infected', main = paste('df =', jj), col = 'gray')
#   lines(min(test.dat$age):max(test.dat$age), plot.probs$fit, col = 'dodgerblue')
#   }
#   ## Continue cross validation
#   sim.outcomes = sapply(X = pred.probs, FUN = function(pp) sample(c(0,1), size = 1, prob = c(1-pp, pp)))
#   test.error.rate[ii, jj] = sum(abs(sim.outcomes - test.dat$prolonged.symp))/nrow(test.dat)
# }
# }
```


Plot the overall test error rate vs. df
```{r echo = FALSE}
# par(mfrow = c(1,1))
# plot(1:max.df, colMeans(test.error.rate), ylab = 'test error rate', xlab = 'df', type = 'b', main = 'H3N2 10-fold CV to choose df')
```

Conclusion: No clear support for a specific # of df. Visually, it looks like 4 df is enough to capture the overall shape with appropriate smoothness for both H1N1 and H3N2.

## Set up a logistic regression that includes the following predictors:
* age
* imprinting status
* vaccination status
* antiviral use
* season
* country





### Clean data for fitting
```{r echo = FALSE}
H1N1.model.dat = dat.002[dat.002$H1N1 == 1 & !is.na(dat.002$prolonged.symp) & !is.na(dat.002$H1N1), ]

# Group countries with 20 or fewer valid cases into "other"
table(H1N1.model.dat$country)
H1N1.model.dat$country[which(H1N1.model.dat$country %in% c('Austria', 'Chile', 'Japan', 'UK'))] = 'Other'

# Convert blocking variables to factors
H1N1.model.dat$anyav = factor(H1N1.model.dat$anyav)
H1N1.model.dat$anyvac = factor(H1N1.model.dat$anyvac)
H1N1.model.dat$anydx = factor(H1N1.model.dat$anydx)
H1N1.model.dat$country = factor(H1N1.model.dat$country)
H1N1.model.dat$season = factor(H1N1.model.dat$season)

# Split into training and test data
nrow(H1N1.model.dat)
test.inds = sample(1:nrow(H1N1.model.dat), size = 840, replace = FALSE)
train.H1N1 = H1N1.model.dat[-test.inds, ]
test.H1N1 = H1N1.model.dat[test.inds, ]
```


### First fit a model where the only independent variable is age
Treat vaccination, antivial use, underlying symptoms, country and season as blocking variables with fixed effects
```{r}
library(ROCR)
library(splines)
## H1N1 fit to country, season and medical history only (no effect of age or imprinting)
fit0 = glm(prolonged.symp ~ anyvac + anyav + anydx + country + season, data = train.H1N1, family = binomial)
summary(fit0)

## H1N1 fit to age only, plus fixed effects of vaccination, av use, underlying conditions, country and season
fit1 = glm(prolonged.symp ~ ns(age, knots = 65) + anyvac + anyav + anydx + country + season, data = train.H1N1, family = binomial)
summary(fit1)

## H1N1 fit to age plus imprinting, plus fixed effects of vaccination, av use, underlying conditions, country and season
fit2 = glm(prolonged.symp ~ ns(age, knots = 65) + p.g1.protection + anyvac + anyav + anydx + country + season, data = train.H1N1, family = binomial)
summary(fit2)


## Plot the fits for each model side by side
par(mfrow = c(1, 2), las = 3, mar = c(6, 6, 6, 3), cex.axis = .7)
library(gam)
plot.gam(fit0, se = T, col = 'dodgerblue', main = 'Baseline')
plot.gam(fit1, se = T, col = 'dodgerblue', main = 'Baseline + age')
plot.gam(fit2, se = T, col = 'dodgerblue', main = 'Baseline + age + imp')
pdf('002Duration_H1N1.pdf')
plot.gam(fit2, se = T, col = 'dodgerblue', main = 'Baseline + age + imp')
dev.off()

## Test performance on test data
# Predict probs for each patient
probs0 = predict(fit0, newdata = test.H1N1, type = 'response')
probs1 = predict(fit1, newdata = test.H1N1, type = 'response')
probs2 = predict(fit2, newdata = test.H1N1, type = 'response')

# Create a predict object
pr0 = prediction(probs0, test.H1N1$prolonged.symp)
pr1 = prediction(probs1, test.H1N1$prolonged.symp)
pr2 = prediction(probs2, test.H1N1$prolonged.symp)

# Assess tp and fp
prf0 = performance(pr0, measure = 'tpr', x.measure = 'fpr')
prf1 = performance(pr1, measure = 'tpr', x.measure = 'fpr')
prf2 = performance(pr2, measure = 'tpr', x.measure = 'fpr')

# Calculate AUCs
auc = sapply(list(pr0, pr1, pr2), FUN = function(xx){aa = performance(xx, measure = 'auc')
                                                                      round(aa@y.values[[1]], 3) })

# Plot AUC
par(mfrow = c(1,1))
plot(prf0, lwd = 2, col = cols[3])
plot(prf1, col = cols[2], add = TRUE, lwd = 2)
plot(prf2, col = cols[1], add = TRUE, lwd = 2)
abline(a = 0, b = 1, lty = 2)
legend('bottomright', paste(c('baseline                  AUC = ', 'baseline+age          AUC = ', 'baseline+age+imp  AUC = '), auc), col = cols[3:1], lty = 1)

pdf('002Duration_AUC_H1N1.pdf')
par(lwd = 3.5, cex = 1.5, cex.axis = 1.5, cex.lab = 1.5)
# Plot AUC
par(mfrow = c(1,1))
plot(prf0, col = cols[3])
plot(prf1, col = cols[2], add = TRUE)
plot(prf2, col = cols[1], add = TRUE)
abline(a = 0, b = 1)
legend('bottomright', paste(c('baseline                  AUC = ', 'baseline+age          AUC = ', 'baseline+age+imp  AUC = '), auc), col = cols[3:1], lty = 1, bty = 'n', cex = 1.5)
dev.off()


AIC = c(fit0$aic, fit1$aic, fit2$aic); names(AIC) =  c('baseline', 'baeline+age', 'baseline+age+imp')
del.AIC = sort(AIC - min(AIC))
del.AIC

load('master.AIC.table.H1N1.RData')
master.AIC.table['002Prolonged',  c('baseline', 'baseline+age', 'baseline+age+imp')] = AIC - min(AIC)
save(master.AIC.table, file = 'master.AIC.table.H1N1.RData')
rm(master.AIC.table)
```




### Clean H3N2 data for fitting
```{r echo = FALSE}
H3N2.model.dat = dat.002[dat.002$H3N2 == 1 & !is.na(dat.002$prolonged.symp) & !is.na(dat.002$H3N2), ]

# Group countries with 20 or fewer valid cases into "other"
table(H3N2.model.dat$country)
H3N2.model.dat$country[which(H3N2.model.dat$country %in% c('Austria', 'Chile', 'Greece', 'Japan', 'Spain', 'UK'))] = 'Other'

# Convert blocking variables to factors
H3N2.model.dat$anyav = factor(H3N2.model.dat$anyav)
H3N2.model.dat$anyvac = factor(H3N2.model.dat$anyvac)
H3N2.model.dat$anydx = factor(H3N2.model.dat$anydx)
H3N2.model.dat$country = factor(H3N2.model.dat$country)
H3N2.model.dat$season = factor(H3N2.model.dat$season)

# Split into training and test data
nrow(H3N2.model.dat)
test.inds = sample(1:nrow(H3N2.model.dat), size = 840, replace = FALSE)
train.H3N2 = H3N2.model.dat[-test.inds, ]
test.H3N2 = H3N2.model.dat[test.inds, ]
```


```{r}
## H3N2 fit to country, season and medical history only (no effect of age or imprinting)
fit0 = glm(prolonged.symp ~ anyvac + anyav + anydx + country + season, data = train.H3N2, family = binomial)
summary(fit0)

## H3N2 fit to age only, plus fixed effects of vaccination, av use, underlying conditions, country and season
fit1 = glm(prolonged.symp ~ ns(age, knots = 65) + anyvac + anyav + anydx + country + season, data = train.H3N2, family = binomial)
summary(fit1)

## H3N2 fit to age plus imprinting, plus fixed effects of vaccination, av use, underlying conditions, country and season
fit2 = glm(prolonged.symp ~ ns(age, knots = 65) + p.g2.protection + anyvac + anyav + anydx + country + season, data = train.H3N2, family = binomial)
summary(fit2)


## Plot the fits for each model side by side
par(mfrow = c(1, 2), las = 3, mar = c(6, 6, 6, 3), cex.axis = .7)
library(gam)
plot.gam(fit0, se = T, col = 'firebrick1', main = 'H3N2 baseline')
plot.gam(fit1, se = T, col = 'firebrick1', main = 'H3N2 baseline + age')
plot.gam(fit2, se = T, col = 'firebrick1', main = 'H3N2 baseline + age + impriting')
pdf('002Duration_H3N2.pdf')
plot.gam(fit2, se = T, col = 'firebrick1', main = 'H3N2 baseline + age + impriting')
dev.off()

## Test performance on test data
# Predict probs for each patient
probs0 = predict(fit0, newdata = test.H3N2, type = 'response')
probs1 = predict(fit1, newdata = test.H3N2, type = 'response')
probs2 = predict(fit2, newdata = test.H3N2, type = 'response')

# Create a predict object
pr0 = prediction(probs0, test.H3N2$prolonged.symp)
pr1 = prediction(probs1, test.H3N2$prolonged.symp)
pr2 = prediction(probs2, test.H3N2$prolonged.symp)

# Assess tp and fp
prf0 = performance(pr0, measure = 'tpr', x.measure = 'fpr')
prf1 = performance(pr1, measure = 'tpr', x.measure = 'fpr')
prf2 = performance(pr2, measure = 'tpr', x.measure = 'fpr')

# Calculate AUCs
auc = sapply(list(pr0, pr1, pr2), FUN = function(xx){aa = performance(xx, measure = 'auc')
                                                                      round(aa@y.values[[1]], 3) })

# Plot AUC
par(mfrow = c(1,1))
plot(prf0, lwd = 2, col = cols[3])
plot(prf1, col = cols[2], add = TRUE, lwd = 2)
plot(prf2, col = cols[1], add = TRUE, lwd = 2)
abline(a = 0, b = 1, lty = 2)
legend('bottomright', paste(c('baseline                  AUC = ', 'baseline+age          AUC = ', 'baseline+age+imp  AUC = '), auc), col = cols[3:1], lty = 1)

pdf('002Duration_AUC_H3N2.pdf')
par(lwd = 3.5, cex = 1.5, cex.axis = 1.5, cex.lab = 1.5)
# Plot AUC
par(mfrow = c(1,1))
plot(prf0, col = cols[3])
plot(prf1, col = cols[2], add = TRUE)
plot(prf2, col = cols[1], add = TRUE)
abline(a = 0, b = 1)
legend('bottomright', paste(c('baseline                  AUC = ', 'baseline+age          AUC = ', 'baseline+age+imp  AUC = '), auc), col = cols[3:1], lty = 1, bty = 'n', cex = 1.5)
dev.off()


AIC = c(fit0$aic, fit1$aic, fit2$aic); names(AIC) =  c('baseline', 'baeline+age', 'baseline+age+imp')
del.AIC = sort(AIC - min(AIC))
del.AIC

load('master.AIC.table.H3N2.RData')
master.AIC.table['002Prolonged',  c('baseline', 'baseline+age', 'baseline+age+imp')] = AIC - min(AIC)
save(master.AIC.table, file = 'master.AIC.table.H3N2.RData')
rm(master.AIC.table)
```

