---
title: "003 hospdays"
author: "Katie Gostic"
date: "11/1/2017"
output: html_document
---

#### Note: hospdays records the number of days hospitalized after study enrollment. In another script, I test tothosp (the total number of hospitalized days). I try to control for hospitalization unrelated to infelunza by excluding patients with underlying conditions from the tothosp analysis

```{r setup, include=FALSE}
setwd('~/Dropbox/R/2017_INSIGHT/')
rm(list = ls())
source('Import_FLU003.R')
set.seed(91)
library(ROCR)
cols = colors()[c(506, 499, 475)]
```
Notes on data formatting from Deborah Wentworth:

* symdur is symptom duration (days) prior to enrollment
* date info is for enrollment date
* fluvac6 and fluvac12 are flu vaccination in last 6 or last 12 months
* anyvac combines values from fluvac6 and fluvac12, whichever was collected for a given patient is reported.
* anyav is use of antivirals at any time
* anydx is any diagnosis in the medical history section
* flutype is subtype: 1=H1N1, 2=H3N2, 3=flu B, 4=PCR negative, 5=Inf A, subtype unknown, and 6=coinfection with multiple strains
* resolved is symptom resolution within 14 days
* country code is the 3 letter code for country of enrollment - think the abbreviations will be obvious




#### Visualize the relationship between age and duration of hospitalization
```{r warning = FALSE, echo = FALSE}
# Select only cases with confirmed H1N1
valid = dat.003[dat.003$H1N1 == 1 & ! is.na(dat.003$H1N1), ]
plot(jitter(valid$age), jitter(valid$hospdays), ylim = c(1, 70), xlim = c(10, 100), xlab = 'age', ylab = 'symdur', col = 'dodgerblue', main = 'H1N1: Duration of hospitalization by age', cex = .7)
cor.test(valid$age, valid$hospdays, method = 'spearman')
```

There is a positive correlation between age and duration of hospitalization.


#### Visualize the relationship between age and prob of H3N2 infection
```{r warning = FALSE, echo = FALSE}
valid = dat.003[dat.003$H3N2 == 1 & !is.na(dat.003$H3N2), ]
plot(jitter(valid$age), jitter(valid$hospdays), ylim = c(0, 70), xlim = c(10, 100), xlab = 'age', ylab = 'symdur', col = 'firebrick1', main = 'H3N2: Duration of hospitalization by age', cex = .7)
cor.test(valid$age, valid$hospdays, method = 'spearman')
```

Results show a positive correlation between duration of hosp and age.

#### Visualize the relationship between country and symptom duration
H1N1

```{r warning = FALSE, echo = FALSE}
valid = dat.003[dat.003$H1N1 == 1 & !is.na(dat.003$H1N1), ]
plot(factor(valid$country), jitter(valid$hospdays),  xlab = '', ylab = 'symdur', col = 'lightblue', main = 'H1N1: Duration of hospitalization by country', cex = .7, las = 3)
```

H3N2

```{r warning = FALSE, echo = FALSE}
valid = dat.003[dat.003$H3N2 == 1 & !is.na(dat.003$H3N2), ]
plot(factor(valid$country), jitter(valid$hospdays),  xlab = '', ylim = c(0, 60), ylab = 'hosp days', col = 'pink3', main = 'H3N2: duration of hospitalization by country', cex = .7, las = 3)
```

# No clear effect of country

#### Visualize the relationship between season and symptom duration

H1N1

```{r warning = FALSE, echo = FALSE}
valid = dat.003[dat.003$H1N1 == 1 & !is.na(dat.003$H1N1), ]
plot(factor(valid$season), jitter(valid$hospdays),  xlab = '', ylab = 'hospital days', col = 'lightblue', main = 'H1N1: duration of hospitalization by season', cex = .7, las = 3)
```

No clear effect of season.



H3N2

```{r warning = FALSE, echo = FALSE}
valid = dat.003[dat.003$H3N2 == 1 & !is.na(dat.003$H3N2), ]
plot(factor(valid$season), jitter(valid$hospdays),  xlab = '', ylim = c(0, 60), ylab = 'hosp days', col = 'pink3', main = 'H3N2: duraiton of hospitalization by season', cex = .7, las = 3)
```


No clear effect of season
--> Try models that don't include season and country as blocking vars




Plot the overall test error rate vs. df
```{r echo = FALSE}
# par(mfrow = c(1,1))
# plot(1:max.df, colMeans(MSE), ylab = 'MSE', xlab = 'df', type = 'b', main = 'H3N2 10-fold CV to choose df')
```

Conclusion: Clear preference for linear fit for H3N2

## Set up a logistic regression that includes the following predictors:
* age
* imprinting status
* vaccination status
* antiviral use
* season
* country




### Clean data for fitting
```{r echo = FALSE}
H1N1.model.dat = dat.003[dat.003$H1N1 == 1 & !is.na(dat.003$anyicu) & !is.na(dat.003$H1N1), ]

# Group countries with 20 or fewer valid cases into "other"
table(H1N1.model.dat$country)
H1N1.model.dat$country[which(H1N1.model.dat$country %in% c('Austria', 'Chile', 'China', 'Norway', 'Peru', 'Poland', 'Singapore'))] = 'Other'

# Convert blocking variables to factors
H1N1.model.dat$anyav = factor(H1N1.model.dat$anyav)
H1N1.model.dat$anyvac = factor(H1N1.model.dat$anyvac)
H1N1.model.dat$anydx = factor(H1N1.model.dat$anydx)
H1N1.model.dat$country = factor(H1N1.model.dat$country)
H1N1.model.dat$season = factor(H1N1.model.dat$season)

# Split into training and test data
nrow(H1N1.model.dat)
test.inds = sample(1:nrow(H1N1.model.dat), size = 77, replace = FALSE)
train.H1N1 = H1N1.model.dat[-test.inds, ]
test.H1N1 = H1N1.model.dat[test.inds, ]


# H3N2
H3N2.model.dat = dat.003[dat.003$H3N2 == 1 & !is.na(dat.003$anyicu) & !is.na(dat.003$H3N2), ]

# Group countries with 20 or fewer valid cases into "other"
table(H3N2.model.dat$country)
H3N2.model.dat$country[which(H3N2.model.dat$country %in% c('Belgium', 'China', 'Peru'))] = 'Other'

# Convert blocking variables to factors
H3N2.model.dat$anyav = factor(H3N2.model.dat$anyav)
H3N2.model.dat$anyvac = factor(H3N2.model.dat$anyvac)
H3N2.model.dat$anydx = factor(H3N2.model.dat$anydx)
H3N2.model.dat$country = factor(H3N2.model.dat$country)
H3N2.model.dat$season = factor(H3N2.model.dat$season)

# Split into training and test data
nrow(H3N2.model.dat)
test.inds = sample(1:nrow(H3N2.model.dat), size = 79, replace = FALSE)
train.H3N2 = H3N2.model.dat[-test.inds, ]
test.H3N2 = H3N2.model.dat[test.inds, ]



# Concatenate H3N2 and H1N1 data
test = rbind(test.H1N1, test.H3N2)
test$H1.valid = test$H3.valid = 0
test$H1.valid[1:nrow(test.H1N1)] = 1
test$H3.valid[(nrow(test.H1N1)+1):nrow(test)] = 1
test$H1.valid = factor(test$H1.valid)
test$H3.valid = factor(test$H3.valid)

train = rbind(train.H1N1, train.H3N2)
train$H1.valid = train$H3.valid = 0
train$H1.valid[1:nrow(train.H1N1)] = 1
train$H3.valid[(nrow(train.H1N1)+1):nrow(train)] = 1
train$H1.valid = factor(train$H1.valid)
train$H3.valid = factor(train$H3.valid)

# Create a master outcome column
test$infected = c(test.H1N1$H1N1, test.H3N2$H3N2)
train$infected = c(train.H1N1$H1N1, train.H3N2$H3N2)

# Create a master protection column
test$p.protected = c(test.H1N1$p.g1.protection, test.H3N2$p.g2.protection)
train$p.protected = c(train.H1N1$p.g1.protection, train.H3N2$p.g2.protection)

```


### First fit a model where the only independent variable is age
Treat vaccination, antivial use, underlying symptoms, country and season as blocking variables with fixed effects
```{r}
library(splines)
## H1N1 fit to medical history only (no effect of country, season, age or imprinting)
fit00 = glm(hospdays ~ anyav + anydx  + H3.valid*anyvac, data = train)
summary(fit00)

## H1N1 fit to country, season and medical history only (no effect of age or imprinting)  
fit0 = glm(hospdays ~ anyav + anydx + country + H3.valid*season + H3.valid*anyvac, data = train)
summary(fit0)

## H1N1 fit to age only, plus fixed effects of vaccination, av use, underlying conditions, country and season
fit1 = glm(hospdays ~ ns(age, knots = 65) + anyav + anydx + country + H3.valid*season + H3.valid*anyvac, data = train)
summary(fit1)

## H1N1 fit to age plus imprinting, plus fixed effects of vaccination, av use, underlying conditions, country and season
fit2 = glm(hospdays ~ ns(age, knots = 65) +p.protected + anyav + anydx + country + H3.valid*season + H3.valid*anyvac, data = train)
summary(fit2)

## H1N1 fit to age only, plus fixed effects of vaccination, av use, underlying conditions NO COUNTRY AND SEASON
fit3 = glm(hospdays ~ ns(age, knots = 65) + anyav + anydx  + H3.valid*anyvac, data = train)
summary(fit3)

## H1N1 fit to age plus imprinting, plus fixed effects of vaccination, av use, underlying conditions, NO COUNTRY AND SEASON
fit4 = glm(hospdays ~ ns(age, knots = 65) + p.protected + anyav + anydx + H3.valid*anyvac, data = train)
summary(fit4)


## Plot the fits for each model side by side
par(mfrow = c(1, 2), las = 3, mar = c(6, 6, 6, 3), cex.axis = .7)
library(gam)
plot.gam(fit00, se = T, col = 'darkslateblue', main = 'Simplified Baseline')
plot.gam(fit0, se = T, col = 'darkslateblue', main = 'Baseline')
plot.gam(fit1, se = T, col = 'darkslateblue', main = 'Baseline + age + imp')
plot.gam(fit2, se = T, col = 'darkslateblue', main = 'Baseline + age + imp')
plot.gam(fit3, se = T, col = 'darkslateblue', main = 'Simple baseline + age + imp')
plot.gam(fit4, se = T, col = 'darkslateblue', main = 'Simple baseline + age + imp')

pdf('003hospdays_constrained.pdf')
plot.gam(fit4, se = T, col = 'darkslateblue', main = 'Simple baseline + age + imp')
dev.off()

# Predict probs for each patient
pred00 = predict(fit00, newdata = test)
pred0 = predict(fit0, newdata = test)
pred1 = predict(fit1, newdata = test)
pred2 = predict(fit2, newdata = test)
pred3 = predict(fit3, newdata = test)
pred4 = predict(fit4, newdata = test)
  
# Estimate the simulated error rate
MSE = numeric(6); names(MSE) = c('simple.baseline', 'baseline', 'baeline+age', 'baseline+age+imp', 'simple.baseline+age', 'simple.baseline+age+imp')
MSE[1] = mean((pred00 - test$hospdays)^2, na.rm = TRUE)
MSE[2] = mean((pred0 - test$hospdays)^2, na.rm = TRUE)
MSE[3] = mean((pred1 - test$hospdays)^2, na.rm = TRUE)
MSE[4] = mean((pred2 - test$hospdays)^2, na.rm = TRUE)
MSE[5] = mean((pred3 - test$hospdays)^2, na.rm = TRUE)
MSE[6] = mean((pred4 - test$hospdays)^2, na.rm = TRUE)
sort(MSE)

AIC = c(fit00$aic, fit0$aic, fit1$aic, fit2$aic, fit3$aic, fit4$aic); names(AIC) =  c('simple.baseline', 'baseline', 'baeline+age', 'baseline+age+imp', 'simple.baseline+age', 'simple.baseline+age+imp')
del.AIC = sort(AIC - min(AIC))
del.AIC

load('master.AIC.table.constrained.RData')
master.AIC.table['003Hospdays',  c('simple.baseline', 'baseline', 'baseline+age', 'baseline+age+imp', 'simple.baseline+age', 'simple.baseline+age+imp')] = AIC - min(AIC)
save(master.AIC.table, file = 'master.AIC.table.constrained.RData')
rm(master.AIC.table)
```




